---
title: "Lab 03 - Nobel laureates"
author: "Adam Paul"
date: "2/11"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

# Exercises

## Exercise 1

### Question 1

How many observations and how many variables are in the dataset? Use inline code to answer this question. What does each row represent?

```{r}
view(nobel)
```


935 observations of 26 variables. Each row represents an individual nobel laureate.

### Question 2

 Create a new data frame called nobel_living that filters for:

1) Laureates for whom country is available

2) Laureates who are people as opposed to organizations (organizations are denoted with "org" as their gender)

3) Laureates who are still alive (their died_date is NA)


```{r G}
nobel_living <- nobel %>% #Is creating the new data frame out of nobel
  filter(!is.na(country), #Is opening filter, and beginning by filtering for ones where country is not NA
  gender != "org",        #Is filtering for ones where gender is not (! meaning not) equal to org
  is.na(died_date))       #Is filtering for ones that do not have a death date
view(nobel_living)
  
```

### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
